---
title: "Effects of the FDI on the Economic Growth of Uganda Time series Evidence from 1960 to 2020"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

The Analysis bases on 4 variables

-   \- FDI stock (look at this a stock of FDI belonging to Uganda proxied with net FDI)

-   \- FDI Inflow

-   \- FDI Transactions (proxying it using REER)

-   \- FDI other Changes (Proxying it using net Foreign Assets)

-   \- GDP of Uganda (This is Yearly)

The time period for analysis is 1960 to 2020, with this raw data having
missing values we need to correct for the missing values statistically.

Then raw data is in USD and Net foreign assets is in LCUs (local
currency Units) which is converted to USD to make it uniform 

The data is cleaned and missing values replaced and the transformation is done in Excel
to produce the dataset used in further analysis here 

-  \- Excel work book is [here](https://github.com/michaelgobz/Final-Research-Project-/blob/main/data.xlsx)

**Brief About the Data in Excel File**
A little detail about the variables contained here 

-   \- FDI stock  (millions of current USD)

-   \- FDI Inflow (millions of current USD)

-   \- REER (Real Effective Exchange Rate)[millions of current USD]

-   \- Excess of NFA (millions of current USD )

-   \- GDP of Uganda (millions of current USD)

The workbook has 6 sheets where sheet 1 is the raw data obtained from the World Bank.
the next 4 sheets include a series of data cleaning and filling in missing values by interpolation, 
and using the HP filter and try to approximate the values of the missing values. this was necessary for further analysis

Sheet 6 of the workbook contains the final clenaed data ready for further analysis. 

**Setting The Analysis Environment**

Setup the packages required for analysis and load them requires the
Rtools to build some of the packages download the binaries from here
this is only for local setup
[Rtools](https://cran.r-project.org/bin/windows/Rtools/rtools42/rtools.html)

**Install the packages**

```{r install packages}
install.packages("AER")
install.packages("dynlm")
install.packages("forecast")
install.packages("readxl")
install.packages("stargazer")
install.packages("scales")
install.packages("quantmod")
install.packages("urca")

```

**Load Installed Packages**
```{r load the packages}
library(AER)
library(dynlm)
library(forecast)
library(readxl)
library(stargazer)
library(scales)
library(quantmod)
library(urca)
library(stats)

```

**Load the data in from Excel*

```{r load the data in data.xlsx}
dataset <- read_xlsx("./data.xlsx",sheet = 6 )
View(dataset)
```

Helper functions for Transformations

```{r helper functions}
# compute BIC for AR model objects of class 'dynlm'
BIC <- function(model) {
  ssr <- sum(model$residualsˆ2)
  t <- length(model$residuals)
  npar <- length(model$coef)
  return(
    round(c("p" = npar - 1,
      "BIC" = log(ssr/t) + (npar + 1) * log(t)/t,
      "R2" = summary(model)$r.squared), 4)
        )
}

AIC <- function(model) {
  ssr <- sum(model$residualsˆ2)
  t <- length(model$residuals)
  npar <- length(model$coef)
  return(
    round(c("p" = npar - 1,
      "AIC" = log(ssr/t) + (npar + 1) * 2/t,
      "R2" = summary(model)$r.squared), 4)
        )
}

# compute logarithms, annual growth rates and 1st lag of growth rates
quants <- function(series) {
    s <- series
      return(
        data.frame("Level" = s,
            "Logarithm" = log(s),
            "AnnualGrowthRate" = 100 * log(s / lag(s)),
            "1stLagAnnualGrowthRate" = lag(100 * log(s / lag(s))))
      )
}


```

**Plots of the variables with x-lab being 1960 to 2020 periods**

```{r Plots the variables}
#plot the lagged log values
GDP <- xts(dataset$`GDP$millions`,order.by = as.Date(dataset$FY, origin = "1899-12-30" ,tryFormats=c("%Y-%m-%d")))
GDP_Growth <- xts(100 * log(GDP/lag(GDP)))
GDP_Growth
plot(as.zoo(GDP_Growth),
      col = "steelblue",
      lwd = 2,
      ylab =  "Growth",
      xlab = "Period in years",
      main = "Ugandas GDP Growth"
)
```

**FDI Inflow**

```{r FDI Inflow }
# FDI inflow 

FDI_inflow <- xts(dataset$`FDIinflows$millions`,order.by = as.Date(dataset$FY, origin = "1899-12-30" ,tryFormats=c("%Y-%m-%d")))
FDI_Growth <- xts(100 * log(FDI_inflow/lag(FDI_inflow)))
View(FDI_Growth)
FDI_Growth
plot(as.zoo(FDI_Growth),
      col = "steelblue",
      lwd = 2,
      ylab = "Growth",
      xlab = "Period in years",
      main = "Ugandas FDI Inflow Growth"
)
```

```{r FDI Stock }
# FDI Stock 

FDI_Stock <- xts(dataset$`FDIStock$millions`,order.by = as.Date(dataset$FY, origin = "1899-12-30" ,tryFormats=c("%Y-%m-%d")))
FDI_Stock_Growth <- xts(100 * log(FDI_Stock/lag(FDI_Stock)))
View(FDI_Stock_Growth)
FDI_Stock_Growth
plot(as.zoo(FDI_Stock_Growth),
      col = "steelblue",
      lwd = 2,
      ylab = "Growth",
      xlab = "Period in years",
      main = "Ugandas FDI Stock Growth"
)
```

**Excess of NFA (net Foreign Assets)**

```{r Excess of NFA }
# Excess of NFA 

NFA_d <- xts(dataset$`ExcessNFA$millions`,order.by = as.Date(dataset$FY, origin = "1899-12-30" ,tryFormats=c("%Y-%m-%d")))
NFA_Growth <- xts(100 * log(NFA_d/lag(NFA_d)))
View(NFA_Growth)
NFA_Growth
plot(as.zoo(NFA_Growth),
      col = "steelblue",
      lwd = 2,
      ylab = "Growth",
      xlab = "Period in years",
      main = "Ugandas Excess of NFA over GDP Growth"
)
```

**REER (Real Effective Exchange Rate)**

```{r REER }
# REER 

REER<- xts(dataset$REER,order.by = as.Date(dataset$FY, origin = "1899-12-30" ,tryFormats=c("%Y-%m-%d")))
REER_Growth <- xts(100 * log(REER/lag(REER)))
View(REER_Growth)
REER_Growth
plot(as.zoo(REER_Growth),
      col = "steelblue",
      lwd = 2,
      ylab = "Growth",
      xlab = "Period in years",
      main = "Ugandas REER Growth"
)
```

**First level of Analysis** At this level of analysis, the
autocorrelation between the different series shall be established then
provide the narrative for the results

\- GDP - FDI_Inflow - FDI_stock - NFA_Spread - REER

**GDP**

```{r GDP ACF and PACF analysis}
#the data set
# GDP

N <- length(GDP_Growth)
N

GDP_level <- as.numeric(GDP_Growth[-1])
GDP_lags <-as.numeric(GDP_Growth[-N])

#estimate model
GDP_mod <- lm(GDP_level ~ GDP_lags)
GDP_armod <- dynlm(ts(GDP_level) ~ L(ts(GDP_level)) + L(ts(GDP_level), 2) + L(ts(GDP_level), 3) + L(ts(GDP_level), 4) )

#robust summary
coeftest(GDP_armod, vcov. = sandwich)
coeftest(GDP_mod, vcov. = sandwich)
#AR(2)
summary(GDP_armod)$sigma
summary(GDP_armod)$r.squared
#AR(1) like ols
summary(GDP_mod)$sigma
summary(GDP_mod)$r.squared

# Calculate the ACF and PACF 
acf(na.omit(GDP_Growth), lag.max = 20, plot= T)
pacf(na.omit(GDP_Growth),lag.max = 20, plot= T)

# Use BIC to determine the lags
BIC(dynlm(ts(GDP_level) ~ 1))

ssr <- sum(GDP_armod$residualsˆ2)
t <- length(GDP_armod$residuals)
npar <- length(GDP_armod$coef)
ssr
t
npar
AIC(GDP_armod)
AICs <- sapply(1:6,FUN=function(x)"AR"=AIC(dynlm(ts(GDP_level) ~ L(ts(GDP_level),1:x))))
AICs

```

**FDI_inflow**

```{r FDI_inflow ACF and PACF analysis}
#the data sets
# FDI_inflow

N <- length(FDI_Growth)
N

FDI_inflow_level <- as.numeric(FDI_Growth[-1])
FDI_inflow_lags <-as.numeric(FDI_Growth[-N])

#estimate model
FDI_inflow_mod <- lm(FDI_inflow_level ~ FDI_inflow_lags)
FDI_inflow_armod <- dynlm(ts(FDI_inflow_level) ~ L(ts(FDI_inflow_level)) + L(ts(FDI_inflow_level), 2) + L(ts(FDI_inflow_level), 3) + L(ts(FDI_inflow_level), 4))

#robust summary
coeftest(FDI_inflow_armod, vcov. = sandwich)
coeftest(FDI_inflow_mod, vcov. = sandwich)
#AR(2)
summary(FDI_inflow_armod)$sigma
summary(FDI_inflow_armod)$r.squared
#AR(1) like ols
summary(FDI_inflow_mod)$sigma
summary(FDI_inflow_mod)$r.squared

# Calculate the ACF and PACF 
acf(na.omit(FDI_Growth), lag.max = 20, plot= T)
pacf(na.omit(FDI_Growth),lag.max = 20, plot= T,main="FDI Inflow")

# Use BIC to determine the lags
BIC(dynlm(ts(FDI_inflow_level) ~ 1))

ssr <- sum(FDI_inflow_armod$residualsˆ2)
t <- length(FDI_inflow_armod$residuals)
npar <- length(FDI_inflow_armod$coef)
ssr
t
npar
AIC(FDI_inflow_armod)
AICs <- sapply(1:6,FUN=function(x)"AR"=AIC(dynlm(ts(FDI_inflow_level) ~ L(ts(FDI_inflow_level),1:x))))
AICs

```

**FDI_stock**

```{r FDI_stock ACF and PACF analysis}
#the data sets
# REER
# NFA_d
# GDP
# FDI_inflow
# FDI_Stock
N <- length(FDI_Stock_Growth)
N

FDI_stock_level <- as.numeric(FDI_Stock_Growth[-1])
FDI_stock_lags <-as.numeric(FDI_Stock_Growth[-N])

#estimate model
FDI_stock_mod <- lm(FDI_stock_level ~ FDI_stock_lags)
FDI_stock_armod <- dynlm(ts(FDI_stock_level) ~ L(ts(FDI_stock_level)) + L(ts(FDI_stock_level), 2) + L(ts(FDI_stock_level), 3) + L(ts(FDI_stock_level), 4))

#robust summary
coeftest(FDI_stock_armod, vcov. = sandwich)
coeftest(FDI_stock_mod, vcov. = sandwich)
#AR(2)
summary(FDI_stock_armod)$sigma
summary(FDI_stock_armod)$r.squared
#AR(1) like ols
summary(FDI_stock_mod)$sigma
summary(FDI_stock_mod)$r.squared

# Calculate the ACF and PACF 
acf(na.omit(FDI_Stock_Growth), lag.max = 20, plot= T)
pacf(na.omit(FDI_Stock_Growth),lag.max = 20, plot= T)

# Use BIC to determine the lags
BIC(dynlm(ts(FDI_inflow_level) ~ 1))

ssr <- sum(FDI_stock_armod$residualsˆ2)
t <- length(FDI_stock_armod$residuals)
npar <- length(FDI_stock_armod$coef)
ssr
t
npar
AIC(FDI_stock_armod)
AICs <- sapply(1:6,FUN=function(x)"AR"=AIC(dynlm(ts(FDI_stock_level) ~ L(ts(FDI_stock_level),1:x))))
AICs
```

**NFA**

```{r NFA_spread ACF and PACF analysis}
#the data sets
# NFA_d

N <- length(NFA_Growth)
N
NFA_Growth

NFA_level <- as.numeric(NFA_Growth[-1])
NFA_lags <-as.numeric(NFA_Growth[-N])

#estimate model
NFA_mod <- lm(NFA_level ~ NFA_lags)
NFA_armod <- dynlm(ts(NFA_level) ~ L(ts(NFA_level)) + L(ts(NFA_level), 2) + L(ts(NFA_level), 3) + L(ts(NFA_level), 4))

#robust summary
coeftest(NFA_armod, vcov. = sandwich)
coeftest(NFA_mod, vcov. = sandwich)
#AR(2)
summary(NFA_armod)$sigma
summary(NFA_armod)$r.squared
#AR(1) like ols
summary(NFA_mod)$sigma
summary(NFA_mod)$r.squared

# Calculate the ACF and PACF 
acf(na.omit(NFA_Growth), lag.max = 20, plot= T)
pacf(na.omit(NFA_Growth),lag.max = 20, plot= T)

# Use BIC to determine the lags
BIC(dynlm(ts(NFA_level) ~ 1))

ssr <- sum(NFA_armod$residualsˆ2)
t <- length(NFA_armod$residuals)
npar <- length(NFA_armod$coef)
ssr
t
npar
AIC(NFA_armod)
AICs <- sapply(1:6,FUN=function(x)"AR"=AIC(dynlm(ts(NFA_level) ~ L(ts(NFA_level),1:x))))
AICs
```

**REER**

```{r REER ACF and PACF analysis }
#the data sets
# REER

N <- length(REER_Growth)
N

REER_level <- as.numeric(REER_Growth[-1])
REER_lags <-as.numeric(REER_Growth[-N])

#estimate model
REER_mod <- lm(REER_level ~ REER_lags)
REER_armod <- dynlm(ts(REER_level) ~ L(ts(REER_level)) + L(ts(REER_level), 2) + L(ts(REER_level), 3) + L(ts(REER_level), 4))

#robust summary
coeftest(REER_armod, vcov. = sandwich)
coeftest(REER_mod, vcov. = sandwich)
#AR(2)
summary(REER_armod)$sigma
summary(REER_armod)$r.squared
#AR(1) like ols
summary(REER_mod)$sigma
summary(REER_mod)$r.squared

# Calculate the ACF and PACF 
acf(na.omit(REER_Growth), lag.max = 20, plot= T)
pacf(na.omit(REER_Growth),lag.max = 20, plot= T)

# adf test for unit root 
summary(ur.df(REER,
  type = "trend",
  lags = 2,
  selectlags = "Fixed"))


```

**Testing and modeling using the ADL and testing for Predictability of
the Regressors** At this level of analysis I shall fit a general
regression model with multiple regressors which is an extension of the
classical ADL model but with multiple Predictors. The mode is as
follows, It uses p lags of the dependent variable and **ql** lags of
**l** additional predictors where **l = 1, . . ., k** then applying the
**Granger's Causality Tests** to a certain the predictive power the lags
included in the model to predict the dependent variable Then Determine
the Dynamic Causal Effects of the Regressors.

```{r ADL model and test for stationality and Predicatble powers}
 GDP_adl <- dynlm(ts(GDP_level) ~ L(ts(GDP_level)) + L(ts(FDI_stock_level)) +L(ts(FDI_stock_level),2) + L(ts(FDI_inflow_level)) +L(ts(FDI_inflow_level),2) + L(ts(NFA_level)) +L(ts(NFA_level),2)+ L(ts(REER_level)) + L(ts(REER_level),2))
coeftest(GDP_adl, vcov. = sandwich)

GDP_adl_1 <- dynlm(ts(GDP_level) ~ L(ts(GDP_level)) + L(ts(FDI_stock_level)) + L(ts(FDI_inflow_level)) + L(ts(NFA_level)) + L(ts(REER_level)))
coeftest(GDP_adl, vcov. = sandwich)

print("ADL with multiple Regressors model's sigma R^2")
summary(GDP_adl)$sigma
print("sigma for the lag 1 ADL")
summary(GDP_adl_1)$sigma
summary(GDP_adl)$r.squared
# contrast with the AR(4) model of GDP_Growth
print("AR(4) model's sigma R^2 ")
summary(GDP_armod)$sigma
summary(GDP_armod)$r.squared

#F-test on coefficients of term spread (Granger's Causality Tests) that uses the F-Test
linearHypothesis(GDP_adl,
c("L(ts(FDI_stock_level))=0","L(ts(FDI_stock_level), 2)=0", "L(ts(FDI_inflow_level))=0","L(ts(FDI_inflow_level), 2)=0", "L(ts(NFA_level))=0","L(ts(NFA_level), 2)=0", "L(ts(REER_level))=0","L(ts(REER_level), 2)=0"),
vcov. = sandwich)

linearHypothesis(GDP_adl,
c("L(ts(FDI_inflow_level))=0" ,"L(ts(FDI_inflow_level), 2)=0"),
vcov. = sandwich)
linearHypothesis(GDP_adl,
c("L(ts(FDI_stock_level))=0","L(ts(FDI_stock_level), 2)=0"),
vcov. = sandwich)

linearHypothesis(GDP_adl,
c("L(ts(NFA_level))=0","L(ts(NFA_level), 2)=0"),
vcov. = sandwich)

linearHypothesis(GDP_adl,
c("L(ts(REER_level))=0","L(ts(REER_level), 2)=0"),
vcov. = sandwich)

```

**Fitting and ARIMAX model and Forecast the Default where h = 20** At
this level, an ARIMAX model is fitted on the series to forecast the GDP
based on the exogenous independent. The ARIMAX model is an extension of
Auto regressive Integrated Moving Average (ARIMA) model. The ARIMA model
has three parameters namely, p, d and q, where p is the autoregressive
term, q is the moving average term and d indicated the series is
differenced to make it stationary.

Regressions with non-stationary series are spurious and analyses are not
valid. For this work, we test if there is stationarity by applying the
Augmented Dickey-Fuller (ADF). Visually examining the autocorrelation
function (ACF) and partial autocorrelation function (PACF) combined
patterns should reveal some important features at least inconclusively
which should help to identify the orders of AR and MA parameters of the
model

```{r ARIMAX modeling}
GDP_arimax <- arima (GDP_Growth, order=c(1, 2, 2),
                xreg=NFA_Growth)
GDP_arimax
GDP_arimax$coef

GDP_arimax_inflow <- arima (GDP_Growth, order=c(1, 2, 2),
                xreg=FDI_inflow)
GDP_arimax_inflow
GDP_arimax_inflow$coef

GDP_arimax_stock <- arima (GDP_Growth, order=c(1, 2, 2),
                xreg=FDI_Stock_Growth)
GDP_arimax_stock
GDP_arimax_stock$coef

REER_arimax <- arima (GDP_Growth, order=c(1, 2, 2),
                xreg=REER_Growth)

REER_arimax
REER_arimax$coef

#values <- forecast(GDP_arimax, newxreg=NFA_Growth)
#values

```

**Narrative**
