---
title: "Effects of the FDI on the Economic Growth of Uganda Time series Evidence"
output: html_notebook
---

The analysis bases on 4 variables

-   \- FDI stock

-    - FDI Inflow

-   \- FDI Transactions (proxying it using REER)

-   \- FDI other Changes (Proxying it using net Foreign Assets)

-   \- GDP of Uganda

The time period for analysis is 1960 to 2021, with this raw data having missing values we need to correct for the missing values statistically.

Then raw data is in USD and Net foreign assets is in LCUs (local currency Units) which is converted to USD to make it uniform

**Setting The Analysis Environment**

Setup the packages required for analysis and load them

Install them
```{r install packages}
```
Load them into usage

```{r load the packages}
```
**Plots of the variables with x-lab being 62 periods**

```{r Plots the variables}
plot(cars)
```

**First level of Analysis**
At this level of analysis, the autocorrelation between the different series shall be established
then provide the narrative for the results 
