---
title: "Effects of the FDI on the Economic Growth of Uganda Time series Evidence"
output: html_notebook
---

The analysis bases on 4 variables

-   \- FDI stock

-    - FDI Inflow

-   \- FDI Transactions (proxying it using REER)

-   \- FDI other Changes (Proxying it using net Foreign Assets)

-   \- GDP of Uganda

The time period for analysis is 1960 to 2021, with this raw data having missing values we need to correct for the missing values statistically.

Then raw data is in USD and Net foreign assets is in LCUs (local currency Units) which is converted to USD to make it uniform

**Setting The Analysis Environment**

Setup the packages required for analysis and load them

Install them
```{r install packages}
```
Load them into usage

```{r load the packages}
```
Helper functions for Transformations 
```{r helper functions}
```
**Plots of the variables with x-lab being 62 periods**

```{r Plots the variables}
plot(cars)
```

**First level of Analysis**
At this level of analysis, the autocorrelation between the different series shall be established
then provide the narrative for the results 

```{r ACF and PACF analysis of the series}
```
**Testing and modeling using the ADL and testing for Predictability of the Regressors**
At this level of analysis I shall fit a general regression model with multiple regressors which is an extension of the classical ADL model but with multiple Predictors. The mode is as follows, It uses p lags of the dependent variable and **ql** lags of **l** additional predictors where **l = 1, . . ., k** then applying the **Granger’s Causality Tests** to a certain the predictive power the lags included in the model to predict the dependent variable Then Determine the Dynamic Causal Effects of the Regressors.
```{r ADL model and test for stationality and Predicatble powers}
```
**Fitting and ARIMAX model and Forecast the Default where h = 20**








